Abstract Binding Trees

### 2015-10-21 paper.tex
notation
fix bug in App notation
Fix incorrect App rule
Fix typo (unbalanced parens)

\newcommand\MkValence[3]{#1\parallel #2\,.\, #3}
\newcommand\MkArity[2]{\left(#1\right) #2}
\newcommand\IsArity[1]{\IMode{#1}\ \textit{arity}}
\newcommand\IsValence[1]{\IMode{#1}\ \textit{valence}}
\newcommand\IsSort[1]{\IMode{#1}\ \textit{sort}}
\newcommand\Leq[2]{\IMode{#1}\leq\IMode{#2}}

@@

\newcommand\IsOperator[3]{\IMode{#1}\vdash \IMode{#2} : \OMode{#3}}

@@

\newcommand\MV[1]{\textsc{#1}}
\newcommand\App[3]{#1\left\{#2\right\}\left(#2;\; #3\right)}
\newcommand\Dom[1]{\left\vert #1\right\vert}
\newcommand\Fresh[2]{\IMode{#1}\notin\IMode{\Dom{#2}}}

@@

Operators are defined by specifying the fibres of $\int\Operators$ in which
they reside. For instance, consider the lambda calculus with a single sort,
$\SortExp$; we give its signature by asserting the following about its
operators:
\begin{gather*}
  \IsOperator{U}{\lambda}{
    \MkArity{\MkValence{\cdot}{\SortExp}{\SortExp}}{\SortExp})
  }\\
  \IsOperator{U}{\mathsf{ap}}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp},
      \MkValence{\cdot}{\cdot}{\SortExp}
    }{\SortExp}
  }
\end{gather*}

@@

\begin{gather*}
  \IsOperator{U}{\mathsf{decl}}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp},
      \MkValence{\SortExp}{\cdot}{\SortExp}
    }{
      \SortExp
    }
  }\\
  \IsOperator{U,u}{\mathsf{get}\{u\}}{
    \MkArity{
      \cdot
    }{
      \SortExp
    }
  }\\
  \IsOperator{U,u}{\mathsf{set}\{u\}}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp}
    }{
      \SortExp
    }
  }
\end{gather*}

Declaring a new assignable consists in providing an initial value, and an
expression with a free symbol (which shall represent the assignable in scope).
Weakening can be seen as inducing a ``degeneracy map'' on operators, whereas a
renaming $u\mapsto v$ will take $\mathsf{get}\{u\}$ to $\mathsf{get}\{v\}$.

@@

Let the judgment $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{s}$ presuppose
$\IsMetaCtx{\Omega}$, $\IsSymCtx\Upsilon$, $\IsVarCtx\Gamma$ and $\IsSort{s}$,
meaning that $M$ is an abstract binding tree of sort $s$, with metavariables in
$\Omega$, parameters in $\Upsilon$, and variables in $\Gamma$. Then, the syntax
of abstract binding trees is inductively defined in three rules:

\begin{gather*}
  \infer[\textit{var}]{
    \IsAbt{\Omega}{\Upsilon}{\Gamma}{x}{s}
  }{
    \Lookup{\Gamma}{x}{s}
  }\\[6pt]
  %
  \infer[\textit{mvar}]{
    \IsAbt{\Omega}{\Upsilon}{\Gamma}{
      \App{\MV{m}}{u_0,\dots,u_m}{M_0,\dots,M_n}
    }{s}
  }{
    \begin{array}{l}
      \Lookup{\Omega}{\MV{m}}{\MkValence{p_0,\dots,p_m}{q_0,\dots,q_n}{s}}\\
      \HypJ{\Lookup{\Upsilon}{u_i}{p_i}}{\Leq{i}{m}}\\
      \HypJ{
        \IsAbt{\Omega}{\Upsilon}{\Gamma}{M_i}{q_i}
      }{
        \Leq{i}{n}
      }
    \end{array}
  }\\[6pt]
  %
  \infer[\textit{app}]{
    \IsAbt{\Omega}{\Upsilon}{\Gamma}{
      \App{\vartheta}{u_0,\dots,u_m}{M_0,\dots,M_n}
    }{s}
  }{
    \begin{array}{l}
      \IsOperator{\Dom{\Upsilon}}{\vartheta}{
        \MkValence{p_0,\dots,p_m}{q_0,\dots,q_n}{s}
      }\\
      \HypJ{\Lookup{\Upsilon}{u_i}{p_i}}{\Leq{i}{m}}\\
      \HypJ{
        \IsAbt{\Omega}{\Upsilon}{\Gamma}{M_i}{q_i}
      }{
        \Leq{i}{n}
      }
    \end{array}
  }
\end{gather*}

>>>>

\newcommand\MkValence[3]{#1;#2\,.\, #3}
\newcommand\MkBTm[3]{#1;#2\,.\, #3}
\newcommand\MkArity[2]{\left(#1\right) #2}
\newcommand\IsArity[1]{\IMode{#1}\ \textit{arity}}
\newcommand\IsValence[1]{\IMode{#1}\ \textit{valence}}
\newcommand\IsSort[1]{\IMode{#1}\ \textit{sort}}
\newcommand\Leq[2]{\IMode{#1}\leq\IMode{#2}}

@@

\newcommand\IsOperator[3]{\IMode{#1}\Vdash \IMode{#2} : \OMode{#3}}

@@

\newcommand\IsBTm[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}:\IMode{#5}%
}

\newcommand\MV[1]{\textsc{#1}}
\newcommand\MApp[3]{#1\left\{#2\right\}\left(#3\right)}
\newcommand\App[2]{#1\left(#2\right)}
\newcommand\Dom[1]{\left\vert #1\right\vert}
\newcommand\Fresh[2]{\IMode{#1}\notin\IMode{\Dom{#2}}}

@@

Operators are defined by specifying the fibres of $\int\Operators$ in which
they reside. For instance, consider the lambda calculus with a single sort,
$\SortExp$; we give its signature by asserting the following about its
operators:
\begin{gather*}
  \IsOperator{U}{\lambda}{
    \MkArity{\MkValence{\cdot}{\SortExp}{\SortExp}}{\SortExp}
  }\\
  \IsOperator{U}{\mathsf{ap}}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp},
      \MkValence{\cdot}{\cdot}{\SortExp}
    }{\SortExp}
  }
\end{gather*}

@@

\begin{gather*}
  \IsOperator{U}{\mathsf{decl}}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp},
      \MkValence{\SortExp}{\cdot}{\SortExp}
    }{
      \SortExp
    }
  }\\
  \IsOperator{U,u}{\mathsf{get}[u]}{
    \MkArity{
      \cdot
    }{
      \SortExp
    }
  }\\
  \IsOperator{U,u}{\mathsf{set}[u]}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp}
    }{
      \SortExp
    }
  }
\end{gather*}

Declaring a new assignable consists in providing an initial value, and an
expression with a free symbol (which shall represent the assignable in scope).
Weakening can be seen as inducing a ``degeneracy map'' on operators, whereas a
renaming $u\mapsto v$ will take $\mathsf{get}[u]$ to $\mathsf{get}[v]$.

@@

Let the judgment $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{s}$ presuppose
$\IsMetaCtx{\Omega}$, $\IsSymCtx\Upsilon$, $\IsVarCtx\Gamma$ and $\IsSort{s}$,
meaning that $M$ is an abstract binding tree of sort $s$, with metavariables in
$\Omega$, parameters in $\Upsilon$, and variables in $\Gamma$. Let the judgment
$\IsBTm{\Omega}{\Upsilon}{\Gamma}{E}{v}$ presuppose $\IsValence{v}$.  Then, the
syntax of abstract binding trees is inductively defined in four rules:

\begin{gather*}
  \infer[\textit{var}]{
    \IsAbt{\Omega}{\Upsilon}{\Gamma}{x}{s}
  }{
    \Lookup{\Gamma}{x}{s}
  }\\[6pt]
  %
  \infer[\textit{mvar}]{
    \IsAbt{\Omega}{\Upsilon}{\Gamma}{
      \MApp{\MV{m}}{u_0,\dots,u_m}{M_0,\dots,M_n}
    }{s}
  }{
    \begin{array}{l}
      \Lookup{\Omega}{\MV{m}}{\MkValence{p_0,\dots,p_m}{q_0,\dots,q_n}{s}}\\
      \HypJ{\Lookup{\Upsilon}{u_i}{p_i}}{\Leq{i}{m}}\\
      \HypJ{
        \IsAbt{\Omega}{\Upsilon}{\Gamma}{M_i}{q_i}
      }{
        \Leq{i}{n}
      }
    \end{array}
  }\\[6pt]
  %
  \infer[\textit{app}]{
    \IsAbt{\Omega}{\Upsilon}{\Gamma}{
      \App{\vartheta}{E_0,\dots,E_n}
    }{s}
  }{
    \begin{array}{l}
      \IsOperator{\Dom{\Upsilon}}{\vartheta}{
        v_1,\dots,v_n
      }\\
      \HypJ{
        \IsBTm{\Omega}{\Upsilon}{\Gamma}{E_i}{q_i}
      }{
        \Leq{i}{n}
      }
    \end{array}
  }\\[6pt]
  \infer[\textit{abs}]{
    \IsBTm{\Omega}{\Upsilon}{\Gamma}{
      (\MkBTm{\vec{u}}{\vec{x}}{E})
    }{
      (\MkValence{\vec{p}}{\vec{q}}{s})
    }
  }{
    \IsAbt{\Omega}{\Upsilon, \vec{u}:\vec{p}}{\Gamma,\vec{x}:{\vec{q}}}{
      E
    }{
      s
    }
  }
\end{gather*}
